## Reproducible Research: Peer Assessment 1

### Loading and preprocessing the data

```{r, echo=TRUE}
activity<-read.csv("activity.csv" , header=TRUE)
head(activity)
```

```{r, echo=TRUE}
library(lattice)
activity$date<-as.Date(activity$date,format = '%Y-%m-%d')
```

### What is mean total number of steps taken per day?

```{r, echo=TRUE}
TotalStepsByDay <-aggregate(steps~date, data=activity,sum, na.rm=TRUE)
head(TotalStepsByDay)
```

```{r, echo=TRUE}
histogram(TotalStepsByDay$steps, main="Total Number of Steps by Day", xlab="Daily Steps")
```

```{r, echo=FALSE}
meanSteps<- mean(TotalStepsByDay$steps)
medianSteps<-median(TotalStepsByDay$steps)
```

On average, the number of steps taken per day is `r meanSteps` and the median is `r medianSteps` 


### What is the average daily activity pattern?

```{r, echo=TRUE}
AverageStepsByInterval <-aggregate(steps~interval, data=activity, mean, na.rm=TRUE)
plot(AverageStepsByInterval$interval, AverageStepsByInterval$steps, type="l",xlab="Interval",ylab="Average Steps", main="Average Number of Steps",col="blue")
```

```{r, echo=TRUE, results="hide"}
maxSteps<-AverageStepsByInterval$interval[which.max(AverageStepsByInterval$steps)]
```

on average, the 5-minute interval that contains the maximum number of steps is `r maxSteps`

### Imputing missing values

```{r, echo=TRUE}
sum(is.na(activity$steps))
```
We replace the NAs with the average number of steps by interval calculated before on the same interval.

```{r, echo=TRUE}
activityNoNA <- activity
for(i in 1:nrow(activityNoNA)){
  if (is.na(activityNoNA$steps[i])) {
    replace <- AverageStepsByInterval$steps[AverageStepsByInterval$interval == activityNoNA$interval[i]];
    activityNoNA$steps[i] <- replace;
  }
}

sum(is.na(activityNoNA$steps))
```

```{r, echo=TRUE}
TotalStepsByDayNoNA <- aggregate(steps ~ date, data=activityNoNA, sum)
histogram(TotalStepsByDayNoNA$steps, main="Total Number of Steps by Day", xlab="Daily Steps")
```

The total number of steps increased around the mean as a result of replacing the missing values.

```{r, echo=TRUE}
mean(TotalStepsByDayNoNA$steps, na.rm=TRUE)
```

```{r, echo=TRUE}
median(TotalStepsByDayNoNA$steps, na.rm=TRUE)
```
The mean and median have very slightly increased.

### Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
activityNoNA$day <- "Weekday"
activityNoNA$day[weekdays(as.Date(activityNoNA$date), abb=T) %in% c("Sat","Sun")]<-"Weekend"
table(activityNoNA$day)
```

```{r, echo=TRUE}
TotalStepsByIntervalNoNADay <- aggregate(steps ~ interval + day, data=activityNoNA, FUN="mean")
xyplot(steps ~ interval | day, data=TotalStepsByIntervalNoNADay, type="l", grid=T, layout=c(1,2), ylab="Number of Steps", xlab="Interval", main="Average 5-min Activity Intervals on Weekdays vs Weekends")
```